---
title: "THE DOCTRINE PACKAGE"
author: "Karel Couceiro"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
```

## Aim of the package

This package is intended to analyze display a dataset consisting of car accidents from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System. Which, is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. At the moment, this package only provides data for 2013, 2014 and 2015.


```{r add_functions, results='hide', echo=FALSE, warning = 'false', message = 'false'}

suppressMessages(library(magrittr, lib.loc="~/R/win-library/3.4"))
suppressMessages(library(dplyr, lib.loc="~/R/win-library/3.4"))
suppressMessages(library(tidyr,lib.loc="~/R/win-library/3.4"))
suppressMessages(library(maps,lib.loc="~/R/win-library/3.4"))

fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}


make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("accident_%d.csv.bz2", year)
}


fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>% 
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}


fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>% 
    dplyr::group_by(year, MONTH) %>% 
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}

fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)
  
  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}

```



**This package provides the necessary tools to load the dataset into the working directory.**


```{r load_all, echo=TRUE ,results='hide'}

fars_read_years(c(2013,2014,2015))

```



**In addition, we can make a brief summary for each subset**. 

Concretely, we can observe how many car accidents ocurred each month for each present year. We can see that the data types are integers.

```{r summary_data}
knitr::kable(fars_summarize_years(c(2013,2014,2015)))
```




**The major capability of the dataset consists in generating a map for a given desired state at a given year, placing in it for a given year the accidents occured.**

For instance, the following map displays the accidents occurred in Alabama in 2015


```{r plot_stat}

fars_map_state(1,2015)

```


